Mergeable Types

### 2017-02-28 paper/meta.tex
Edits.

\begin{figure}[!t]
\centering
\subcaptionbox[] {\small
  In this example, 1 and 3 have two LCAs (3 and 4) a result of
  previous merges. The dotted circle denotes a virtual ancestor
  obtained by merging the two LCAs.
  \label{fig:criss-cross-lcas}
} [0.47\columnwidth] {
  \includegraphics[scale=0.55]{Figures/2-LCAs}
}
\hfill
\subcaptionbox[] {\small
  In this example, versions $v_{13}$ and $v_{44}$ have two LCAs
  ($v_{22}$ and $v_{32}$)  despite there not being any previous merges
  between their respective branches.
  \label{fig:external-lcas}
} [0.47\columnwidth] {
  \includegraphics[scale=0.7]{Figures/2-external-LCAs}
}
\caption{Examples where two branches have more than one LCA, hence
cannot merge. }
\label{fig:many-lcas}
\end{figure}

@@

Fortunately, unlike source control systems where branching structure
is entirely dictated by the user, \name abstracts away branching
structure from the programmer, and hence retains the ability to
manifest it in a way that it deems fit. In particular, \name solves
the problem of multiple LCAs by suitably constraining the branching
structure such that the problem never arises. The constraints are
imposed either implicitly, as a result of how operational semantics
defines an atomic step, or explicitly, by insisting that certain
conditions be met before merging a pair of versions
(\rulelabel{E-Pull-Wait}). Firstly, the operational semantics already
disables criss-cross merges since it only ever merges versions that
are latest on their respective branches. For instance, suppose
$v_{11}$ and $v_{21}$ are the latest versions on branches $b_1$ and
$b_2$, respectively. Merging $b_2$ into $b_1$ entails merging $v_{21}$
into $v_{11}$ to generate version $v_{12}$ on $b_1$.  Now, merging
$b_1$ into $b_2$ involves merging $v_{12}$, the latest version on
$b_1$, into $v_{21}$, but not $v_{11}$ into $v_{21}$, thus preempting
a criss-cross branching structure.

Secondly, we impose certain pre-conditions on the merging branches to
preempt the structure shown in Fig.~\ref{fig:external-lcas}. The
intuition is as follows: consider the branch $b_3$ at the instance of
merging $b_1$. Since it has already merged $b_2$, a version on $b_2$
($v_{21}$) could be a common ancestor for $b_3$ and some other branch
(call it $b$). Now, if $b_3$ merges $b_1$, same could be true of $b_3$
and $b_1$: a version on $b_1$ ($v_{11}$) could be a common ancestor
for $b_3$ and the other branch $v$. Since $v_{21}$ is also a common
ancestor for both the branches, and the ancestors are not ordered by
the ancestor relation, $b_3$ and $b$ end up with two LCAs. In
Fig.~\ref{fig:external-lcas}, the role of $b$ is played both by $b_1$
and $b_2$. We observe that this scenario can be prevented if, when
merging $b_1$, $b_3$ insists on an ancestor relation between the last
merged version ($v_{22}$) and the currently merging version.  We call
the last merged version of a branch its \emph{locus}. By maintaining a
total order of among locii of a branch $b$ as new versions are
created, we effectively guarantee the presence of a version (the
locus) that is lower than all the ancestors of the latest version on
$b$. Next, by requiring the locus to be an ancestor of the merging
version, we ensure that all the common ancestors have a single lowest
element, thus enforcing the uniqueness of the LCA.

>>>>

\begin{figure}[!t]
\centering
\subcaptionbox[] {\small
  In this example, 1 and 3 have two LCAs (3 and 4) a result of
  previous merges. The dotted circle denotes a virtual ancestor
  obtained by merging the two LCAs.
  \label{fig:criss-cross-lcas}
} [0.47\columnwidth] {
  \includegraphics[scale=0.55]{Figures/2-LCAs}
}
\hfill
\subcaptionbox[] {\small
  In this example, latest versions on $b_1$ and $b_3$, and
  $b_2$ and $b_3$ have two LCAs each, hence are unmergeable. 
  \label{fig:external-lcas}
} [0.47\columnwidth] {
  \includegraphics[scale=0.7]{Figures/2-external-LCAs}
}
\caption{Examples where two branches have more than one LCA, hence
cannot merge. }
\label{fig:many-lcas}
\end{figure}

@@

Fortunately, unlike source control systems where branching structure
is entirely dictated by the user, \name abstracts away branching
structure from the programmer, and hence retains the ability to
manifest it in a way that it deems fit. In particular, \name solves
the problem of multiple LCAs by suitably constraining the branching
structure such that the problem never arises. The constraints are
imposed either implicitly, as a result of how operational semantics
defines an atomic step, or explicitly, by insisting that certain
conditions be met before merging a pair of versions
(\rulelabel{E-Pull-Wait}). Firstly, the operational semantics already
disables criss-cross merges since it only ever merges versions that
are latest on their respective branches. Secondly, we impose certain
pre-conditions on the merging branches to preempt the structure shown
in Fig.~\ref{fig:external-lcas}. The intuition is as follows: consider
the branch $b_3$ at the instance of merging $v_{12}$. Since it has
already merged $v_{22}$, $v_{22}$ could be a common ancestor for $b_3$
and some other branch (call it $b$). Now, if $b_3$ merges $v_{12}$,
same could be true of $v_{12}$. Since $v_{12}$ and $v_{22}$ are not
orderd by the ancestor relation, both become LCAs of
$b_3$ and $b$. We observe that this scenario can be prevented if, when
merging $b_1$, $b_3$ insists on an ancestor relation between the last
merged version ($v_{22}$) and the currently merging version.  We call
the last merged version of a branch its \emph{locus}. By maintaining a
total order of among locii of a branch $b$ as new versions are
created, we effectively guarantee the presence of a version (the
locus) that is lower than all the ancestors of the latest version on
$b$. Next, by requiring the locus to be an ancestor of the merging
version, we ensure that all the common ancestors have a single lowest
element, thus enforcing the uniqueness of the LCA.
